---
title: "ASCI 431/831 Advanced Animal Breeding"
author: "Review of allele and genotypic frequencies"
date: "2/12/2018"
output: html_document
header-includes:
  - \usepackage{color}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### <span style="color:darkblue">Read Santa Ines sheep data</span> 

```{r}
W<-read.table('/Users/gerardo.mamani/Dropbox/2Santa_ines/inbreeding/grm/W.txt', header = F)
#W <- read.table(file = file.choose(), header = TRUE, stringsAsFactors = FALSE)
```

```{r, comment=NA}
W[1, 1]       # 1st row and 1st column 
head(W[, 1]) 
```
```{r, comment=NA}
W[1:5, 1:5]
```

We then drop the first column of data frame, which is the animal IDs. The - sign indicates dropping variables. So, -1 means dropping the first column.
```{r, comment=NA}
W <- W[, -1]
```

Next, we will convert W into a matrix from a data frame. In R, matrices are more memory efficient and convenient than the other data types to do linear algebra.

```{r, comment=NA}
W <- as.matrix(W)
```

What is the dimension of W?
```{r, comment=NA}
dim(W)
```

### <span style="color:darkblue">Allele frequency </span> 
```{r, comment=NA}
table(W[,1])
```

```{r, comment=NA}
nAA <- table(W[, 1])[3]
nAa <- table(W[, 1])[2]
naa <- table(W[, 1])[1]
```

#### <span style="color:darkblue">Exercise 1 </span> 

Use the variables nAA, nAa, and naa defined above and compute the allele frequencies of A and a in the first SNP.

```{r, comment=NA}
fA<-(2*nAA+nAa)/(2*(nrow(W)) )
fA
```
```{r, comment=NA}
fa<-(2*naa+nAa)/(2*(nrow(W)) )
fa
```

### <span style="color:darkblue">Genotypic frequency </span> 

#### <span style="color:darkblue">Exercise 2 </span> 

What are the genotypic frequencies of AA, Aa, and aa in the first SNP?

```{r, comment=NA}
P<-nAA/nrow(W) ;P
H<-nAa/nrow(W) ;H
Q<-naa/nrow(W) ;Q
```

### <span style="color:darkblue">Another approach for obtaining allele frequency</span> 

#### <span style="color:darkblue">Exercise 3 </span> 

Use the variables P, H, and Q defined above and compute the allele frequencies of A and a in the first SNP. 


```{r, message=FALSE, warning=FALSE, comment=NA}
p <- (2*P + 1*H)/(2*(P + H + Q))
p
q <- (2*Q + 1*H)/(2*(P + H + Q))
q
```

#### <span style="color:darkblue">Exercise 4 </span> 

```{r}
nAA <- table(W[, 2])[3]
nAa <- table(W[, 2])[2]
naa <- table(W[, 2])[1]
```
```{r}
p <- (2 * nAA + 1 * nAa)/(2 * (nAA + nAa + naa))
p
q <- 1 - p
q
```

### <span style="color:darkblue">Compute allele frequencies for all SNPs </span> 

Recall that allele frequency of A is given by 

```{r, comment=NA}
sum(W[,1])  # sum of A allele in the first SNP
sum(W[,2])  # sum of A allele in the second SNP
```

```{r, results='hide'}
colSums(W)
```
Note that colSums(W) gives the numerator of the above equation. We then divide this number by 2×total no. of individuals. The function nrows returns the number of rows.

```{r, comment=NA}
p <- colSums(W)/(2 * nrow(W))
```
The variable p is a vector and it contains the allele frequencies of reference allele for 12,703 SNPs.

#### <span style="color:darkblue">Exercise 5 </span>

What is the allele frequency of reference allele in the 300th SNP?

```{r}
p300<-sum(W[,300])/(2*nrow(W))
p300
```

#### <span style="color:darkblue">Exercise 6 </span>

What is the mean of reference allele frequencies in this population?

```{r, comment=NA}
mean(p, na.rm = T)
```
### <span style="color:darkblue">Minor allele frequency </span> 

```{r}
maf<-ifelse(p>0.5, 1-p,p)
```

#### <span style="color:darkblue">Exercise 7 </span> 

```{r, comment=NA}
maf[300]
```

#### <span style="color:darkblue">Exercise 8 </span> 

What is the minor allele frequency of reference allele in the 300th SNP

```{r}
mean(maf, na.rm=TRUE)

```

### <span style="color:darkblue">Visualization </span> 

```{r}
plot(maf[1:500], pch=20)

```

### <span style='color:darkblue'>Save R objects </span>

```{r}
#save(W,file="W.Rda")
```
